Using Distributed Information Stores for Patient-Clinician Health Data Sharing

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
phrs

\section{Introduction}

A key goal in improving patient-clinician communication is
establishing a common context for communication with which a shared
understanding can be established.  This understanding is
bi-directional, in the sense that, both the patient needs to be able
to properly interpret what the clincian is trying to convey to him or
her, while the clinician needs to be able to arrive at an accurate
characterisation of the patient's situation, as updated since the last
time he or she met with the patient.

One of the most promising technologies to help provide a common
context of communication is the patient-centric electronic health
record\cite{ehr} which constitute a complete medical history of
a patient as recorded through his or her entire life. (...)

% challenges to 
% background on authors
%

>>>>

\section{Introduction}
A key goal in improving patient-clinician communication is
establishing a common context for communication with which a shared
understanding can be established.  This understanding is
bi-directional, in the sense that the patient needs to be able to
effectively interpret what the clincian is trying to convey to him or
her, but also that the clinician needs to be able to arrive at an
accurate characterisation of the patient's situation and condition, as
updated since the patient was last seen, often under time pressure and
other constraints.

One of the most promising technologies that has the potential to help
provide such a common context is the patient-centric electronic health
record (henceforth EHR)\cite{ehr}, a digital representation of the
complete medical history of an individual as recorded by clinicians
throughout a patient's life. The EHR has enabled both patients and
their clinicians unfettered access to the patient's medical data,
including diagnostics, reports, and history of interventions.  This
means that clinicians can more easily review a patient's long-term
history even under time pressure and constriants. Moreover, since
patients have access to their own medical histories, they can access
the raw information that was the basis for a prognosis or decision.

% proactive activity recording to provide sketches of a person's
% activity profiles outside of the hospital, streaming data

Yet, so far, EHRs have primarily been limited to the storage of
official diagnoses and diagnostic reports prepared by clinicians,
technicians and nurses.  These communications are both highly
technical, making them difficult to interpret by end-user patients,
but also highly limited in temporal resolution, reflecting only cases
where a person was attended to and diagnostic-tested by a medical
professional.  Although this may be frequent for patients in inpatient
care, those that come to see their clinician, say, once a month end up
with extremely low-resolution views of a patient's condition.

In this position paper, we look at the potential and problems of
augmenting the next generation of EHRs with \emph{personal activity
  diaries}, portraits of a person's dynamic condition captured at high
temporal fidelity.  Such activity diaries, we posit, could be derived
from a combination of automatically-sensed activity and vital
statistics through a wide array of available consumer-grade
network-enabled activity senors and diagnostic tools, as well as
through manual annotations provided by patients about their day.
